### 2025-04-21 sec/5_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we presented a general approach of minimizing the Truncated Least Squares (TLS) objective motivated by the geometry of the problem. We showed how intersection graphs allow for an elegant formulation of the combinatorial problem and how the global minimum can be found among the maximal cliques. We showed that surprisingly, the previously studied compatibility graphs motivated by problem-specific geometric observations lead to these special classes of graphs, namely the intersection graphs. 
We give a sufficient condition under which the global optimum is a maximal clique and discuss cases where subcliques may be the global minimum.
Unfortunately, we were not able to fully prove that we can detect all conditions under which subcliques may be the global optimum.
We think that a better understanding of the $\SEthree$-manifolds and their geometry including their combinatorial properties could lead to efficient algorithms for detecting such cases.