### 2025-04-21 sec/downprojection_lb.tex

\textbf{Down-projecting the lower bound from TIMS-Rot}. After solving the TIMS-Rot problem, we can obtain from it a lower bound on the original registration problem (Eq. \ref{eq:pcr-tls}). We first obtain from (already minimized) WLS relaxation the number of potential inliers $N_{in}$ and the inlier score value $W_{in}$:

\begin{equation}
	\label{eq:wls-relax-info}
	\begin{aligned}
		N_{in} &:= \sum_{i=1}^{N} (1 - o_i)\\
		W_{in} &:= \sum_{i=1}^{N} (1 - o_i) \left(w_i r_i  + (1 - w_i) r_{min}^i \right)
	\end{aligned}
\end{equation}

TODO prove. after theorem 6 turns out to be wrong, do we need to alter this formula? like do we already count each vertex touched by at least one edge?\\
\begin{equation}
	\label{eq:lb-downprojection}
	\begin{aligned}
		R_{in} &= \sqrt{\frac{2 N_{in} + N}{N^2}}\\
		TLS_{in} &= \frac{W_{in}}{N \cdot R_{in}}\\
		TLS_{out} &= (1 - R_{in}) \epsilon^2\\
		project-lb  &= tls_{in} + tls_{out}
	\end{aligned}
\end{equation}