### 2025-04-21 sec/eval-table.tex

\begin{table*}[!ht]
	\centering
	\scriptsize
	\renewcommand\tabcolsep{4.0pt}
	\begin{tabular}{cc|cccccc|cccccc|c}%\item The \LaTeX{}-generated files (e.g. .aux and .bbl file, etc.) if they are needed to compile your source (generally, the .aux and bbl file can be omitted; the bib file might be necessary, however).

		\hline%\end{itemize}
		& & \multicolumn{6}{c|}{FPFH (traditional descriptor)} & \multicolumn{6}{c|}{FCGF (learning based descriptor)} \\ 
		& & RR(\%$\uparrow$) & RE($\circ$$\downarrow$) & TE(cm$\downarrow$) & IP($\%$$\uparrow$) &  IR($\%$$\uparrow$) & F1($\%$$\uparrow$) & RR($\%$$\uparrow$) & RE($\circ$$\downarrow$) & TE(cm$\downarrow$) & IP($\%$$\uparrow$) &  IR($\%$$\uparrow$) & F1($\%$$\uparrow$) & Time(s) \\ \hline
		\multirow{7}{0.04cm}{\rotatebox{90}{learning}} & DCP* \cite{wang2019deep} & 3.22 & 8.42 & 21.40  & - & - & - & 3.22 & 8.42 & 21.40  & - & - & - & 0.07\\
		& PointNetLK* \cite{aoki2019pointnetlk} & 1.61 & 8.04  & 21.30 & - & - & - & 1.61 & 8.04 & 21.30 & - & - & - & 0.12 \\
		& 3DRegNet \cite{pais20203dregnet} & 26.31 & 3.75 & 9.60 & 28.21 & 8.90 & 11.63 & 77.76 & 2.74 & 8.13 & 67.34 & 56.28 & 58.33 & \textcolor{Blue}{0.05}\\
		& DGR \cite{choy2020deep} & 32.84 & 2.45 & 7.53 & 29.51 & 16.78 & 21.35 & 88.85 & 2.28 & 7.02 & 68.51 & 79.92 & 73.15 & 1.53\\

		& DHVR \cite{lee2021deep} & 67.10 & 2.78 & 7.84 & 60.19 & 64.90 & 62.11 & 91.93 & \textcolor{Blue}{2.25} & 7.08 & \textcolor{Blue}{80.20} & 78.15 & 78.98 & 3.92\\
		& DHVR-Origin \cite{lee2021deep} & \textcolor{Blue}{80.22} & \textcolor{Blue}{2.06} & 6.87 & & & & 91.40 & \textcolor{Red}{2.08} & 6.61 & & & & 0.46\\
		& PointDSC \cite{bai2021pointdsc} & 77.57 & \textcolor{Red}{2.03} & \textcolor{Red}{6.38} & 68.45 & \textcolor{Blue}{71.56} & \textcolor{Blue}{69.75} & \textcolor{Blue}{92.85} & \textcolor{Red}{2.08} & \textcolor{Red}{6.51} & 78.91 & 86.23 & \textcolor{Blue}{82.12} & 0.10\\ \hline
		% 		& PointDSC-gen \cite{bai2021pointdsc} & 68.12 & 2.69 & 7.91 & \textcolor{Red}{83.63} & 20.06 & 31.12 & 87.74 & \textcolor{Red}{1.76} & \textcolor{Red}{6.02} & \textcolor{Red}{95.77} & 42.39 & 57.19 & 0.10\\  \hline

		\multirow{10}{0.04cm}{\rotatebox{90}{traditional}} & SM \cite{leordeanu2005spectral} & 55.88 & 2.94 & 8.15 & 47.96 & 70.69 & 50.70 & 86.57 & 2.29 & 7.07 & 81.44 & 38.36 & 48.21 & \textcolor{Red}{0.03}\\
		& ICP* \cite{besl1992method} & 5.79 & 7.93 & 17.59 & - & - & - & 5.79 & 7.93 & 17.59 & - & - & - & 0.25\\
		& FGR \cite{zhou2016fast} & 40.91 & 4.96 & 10.25 & 6.84 & 38.90 & 11.23 & 78.93 & 2.90 & 8.41 & 25.63 & 53.90 & 33.58 & 0.89 \\
		& TEASER \cite{yang2020teaser} & 75.48 & 2.48 & 7.31 & \textcolor{Red}{73.01} & 62.63 & 66.93 & 85.77 & 2.73 & 8.66 & \textcolor{Red}{82.43} & 68.08 & 73.96 & 0.07\\
		& GC-RANSAC \cite{barath2018graph} & 67.65 & 2.33 & 6.87 & 48.55 & 69.38 & 56.78 & 92.05 & 2.33 & 7.11 & 64.46 & \textcolor{Red}{93.39} & 75.69 & 0.55\\
		%& RANSAC-1k & 40.05 & 5.16 & 13.65 & 51.52 & 34.31 & 39.23 & 86.57 & 3.16 & 9.67 & 76.86 & 77.45 & 76.62 & 0.08\\
		& RANSAC-1M \cite{fischler1981random} & 64.20 & 4.05 & 11.35 & 63.96 & 57.90 & 60.13 & 88.42 & 3.05 & 9.42 & 77.96 & 79.86 & 78.55 & 0.97 \\
		& RANSAC-2M \cite{fischler1981random} & 65.25 & 4.07 & 11.56 & 64.41 & 58.37 & 60.51 & 90.88 & 2.71 & 8.31 & 78.52 & 83.52 & 80.68 & 1.63 \\
		& RANSAC-4M \cite{fischler1981random} & 66.10 & 3.95 & 11.03 & 64.27 & 59.10 & 61.02 & 91.44 & 2.69 & 8.38 & 78.88 & 83.88 & 81.04  & 2.86 \\
		& CG-SAC \cite{quan2020compatibility} & 78.00 & 2.40 & 6.89 & 68.07 & 67.32 & 67.52 & 87.52 & 2.42 & 7.66 & 75.32 & 84.61 & 79.90 & 0.27 \\
		& Ours & \textcolor{Red}{83.98} & 2.18 & \textcolor{Blue}{6.56} & \textcolor{Blue}{72.48} & \textcolor{Red}{78.33} & \textcolor{Red}{75.10} & \textcolor{Red}{93.28} & \textcolor{Red}{2.08} & \textcolor{Blue}{6.55} & 78.94 & \textcolor{Blue}{86.39} & \textcolor{Red}{82.20} & 0.11\\ \hline

	\end{tabular}
	\caption{Registration accuracy results on KITTI dataset compared with other methods}
	\vspace{-3mm}
	\label{tab:eval-table}
\end{table*}