### 2025-05-05 sec/5_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we presented a general approach of minimizing the Truncated Least Squares (TLS) objective function 
using a fast linear time convex relaxation and Branch-and Bound. Our convex relaxation is general and can be applied for every model such as translation, rotation and rigid transform as long as residual intervals can be computed. We gave a full proof
for the correctness of the relaxation, which is the main part of the proof of global optimality of a BnB algorithm.
Our evaluation shows that the cost to compute and minimize the relaxation is in the single milliseconds range, allowing to solve the 
translation and rotation-only problem two orders of magnitude faster than with the SoTA SDP-based method STRIDE.
More work is however needed to solve the full 6DoF rigid transformation problem. Altought our relaxation can be applied here as well, branching over the 6D space is overall too slow. Further work in this direction will focus on contracting the rotation search space given a translation space, so that the search is not excessively inefficient.